Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions

### 2023-10-01 latex/intro.tex
Submit to Bayesian Analysis

\section{Introduction}\label{section:intro}
Measurements play a fundamental role in generating observations and
are indispensable to any inference process. In the practice of inverse
problems, selecting the optimal set of measurements holds particular
significance when inferring parameters of a physical process. Unlike
many other fields, where measurements are fixed, in this context, we
have the freedom to choose which measurements to take. This freedom
should be taken advantage of: measurements should be chosen to enhance
accuracy, reduce costs, or both. Whether measurements involve
specifying electrode locations on the skin in electric impedance
tomography \cite{horesh2010impedance}, determining certain wavelengths
in MRI \cite{horesh2008mri}, positioning sensors for detecting
ground-reflected waves in the search for oil
\cite{horesh2008borehole}, or optimizing sensor locations in a
seismographic network \cite{rabinowitz1990, steinberg1995} --- the
selection of optimal measurements, referred to as the problem of
\emph{optimal design}, becomes crucial. Optimal measurements are
chosen according to specific \emph{design criteria}, with A- and
D-optimality being two of the most widely recognized and extensively
studied design criteria \cite{Chaloner1995}.

Surprisingly, A- and D-optimal designs have been observed to yield
remarkably similar measurements in certain cases \cite{fedorov1996,
  nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
illustrate this point, we consider a toy inverse problem: inferring
the initial condition of the 1D heat equation (details in the
supplementary material). In Fig.~\ref{fig:clusterization
  illustration} D-optimal measurement locations are shown for
different numbers of measurements. Notably, for six measurements, a
D-optimal design yields two sets of measurements that are almost
indistinguishable from one another. We refer to this intriguing
phenomenon as \emph{measurement clusterization} \cite{Ucinski05}, and
we refer to a design that exhibits measurement clusterization as a
\emph{clustered design}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{eigenvectors.pdf}
    \caption{D-optimal measurement locations ($m=4$ measurements) and
      weighted eigenvectors for the inverse problem of finding the
      initial condition of the 1D heat equation. Measurement locations
      and weighted eigenvectors are plotted over the computational
      domain $\Omega = [0, 1]$ (x-axis). Measurement clusterization
      occurs approximately at $0.31$ and $0.69$. These two locations
      are a compromise between zeros of eigenvectors a D-optimal
      design aims to ignore (third and up) and staying far from zeros
      of the eigenvectors a D-optimal design aims to measure (first
      and second). Allocating $m=4$ measurements into two locations
      results in clusterization, according to the pigeonhole
      principle.}
  \label{fig:why}
\end{figure}

>>>>

\section{Introduction}\label{section:intro}
Measurements play a fundamental role in generating observations and
are indispensable to any inference process. Selecting the optimal set
of measurements holds particular significance when inferring
parameters of a physical process. Unlike many other fields, where
measurements are fixed, in this context, we have the freedom to choose
which measurements to take. This freedom should be taken advantage of:
measurements should be chosen to enhance accuracy, reduce costs, or
both. Whether measurements involve specifying electrode locations on
the skin in electric impedance tomography \cite{horesh2010impedance},
determining certain wavelengths in MRI \cite{horesh2008mri},
positioning sensors for detecting ground-reflected waves in the search
for oil \cite{horesh2008borehole}, or optimizing sensor locations in a
seismographic network \cite{rabinowitz1990, steinberg1995} --- the
selection of optimal measurements, referred to as the problem of
\emph{optimal design}, becomes crucial. Optimal measurements are
chosen according to specific \emph{design criteria}, with A- and
D-optimality being two of the most widely recognized and extensively
studied design criteria \cite{Chaloner1995,
  AlexanderianGloorGhattas14}.

Surprisingly, A- and D-optimal designs have been observed to yield
remarkably similar measurements in certain cases \cite{fedorov1996,
  nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
illustrate this point, we consider a toy problem: inferring
the initial condition of the 1D heat equation (details in the
supplementary material). In Fig.~\ref{fig:clusterization
  illustration} D-optimal measurement locations are shown for
different numbers of measurements. Notably, for six measurements, a
D-optimal design yields two sets of measurements that are almost
indistinguishable from one another. We refer to this intriguing
phenomenon as \emph{measurement clusterization} \cite{Ucinski05}, and
we refer to a design that exhibits measurement clusterization as a
\emph{clustered design}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{eigenvectors.pdf}
    \caption{D-optimal measurement locations ($m=4$ measurements) and
      weighted eigenvectors for finding the initial condition of
      the 1D heat equation. Measurement locations and weighted
      eigenvectors are plotted over the computational domain $\Omega =
      [0, 1]$ (x-axis). Measurement clusterization occurs
      approximately at $0.31$ and $0.69$. These two locations are a
      compromise between zeros of eigenvectors a D-optimal design aims
      to ignore (third and up) and staying far from zeros of the
      eigenvectors a D-optimal design aims to measure (first and
      second). Allocating $m=4$ measurements into two locations
      results in clusterization, according to the pigeonhole
      principle.}
  \label{fig:why}
\end{figure}

### 2023-10-01 latex/ms.tex
Submit to Bayesian Analysis

\title{Measurement Clusterization in D-optimal Designs for Bayesian
  Linear Inverse Problems over Hilbert Spaces
}

\author[1]{Yair Daon}

\address[1]{Azrieli Faculty of Medicine, Bar Ilan University, Safed,
  Israel}

@@

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}
%% \input{how}
%% \section{Open Problems}

@@

%% \input{appendix}
\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/bibtex.bib}

@@

>>>>

\title[Measurement Clusterization in D-optimal Designs]{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}

\author{Yair Daon}
\address{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel}

@@

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}

@@

\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/bibtex.bib}

@@

%% \input{how}
%% \section{Open Problems}
